---
title: "Homework 3"
author: "Thiago de Araujo - UNI tbd2117"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(ggridges)
library(hexbin)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

Loading Intracart data.

```{r}
library(p8105.datasets)
data("instacart")
```

**Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.**

```{r, message = FALSE}
order_df=
instacart %>% 
  group_by(order_id) %>% 
  summarize(n_item = n())
```

The _instacart_ dataset constains data on products ordered through a online grocery service that allows you to shop online from local stores. The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns. 
Each row in the dataset is a product from an order (there is a single order per user in this dataset).  
It contains products from `r n_distinct(pull(instacart, order_id))` orders with an average `r round(mean(pull(order_df, n_item)), digits = 2)` items per order. 
Products are present in aisles within different departments.

**How many aisles are there, and which aisles are the most items ordered from?**

```{r, message = FALSE}
aisle_df =
  instacart %>% 
    group_by(aisle) %>% 
    summarize(n_item = n()) %>%
    arrange(desc(n_item))
```

```{r, include = FALSE}
#live session
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are `r nrow(aisle_df)` aisles with most itens ordered from the following aisles: `r head(pull(aisle_df, aisle))`.

**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r, message = FALSE}
aisle_df %>% 
  filter(n_item > 10000) %>%
  ggplot(aes(x = reorder(aisle, desc(n_item)), y = n_item)) + 
  geom_col() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle=45, hjust = 1)
  ) + 
  labs(
    title = "Number of Items Ordered in each Aisle",
    x = "Aisle",
    y = "Number of Items Ordered",
    caption = "Limited by aisles with more than 10,000 items orderd"
  )
```

```{r, include = FALSE}
# live session
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = as.factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle=90, hjust = 1)
  )
```

**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r, message = FALSE}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  select(aisle, product_name) %>%
  group_by(aisle, product_name) %>%
  summarise(
    n_item = n()
  ) %>%
  arrange(desc(n_item)) %>%
  slice(1:3) %>% 
  mutate(aisle = str_to_title(aisle)) %>% 
  knitr::kable(
    col.names = c("Aisle", "Product Name", "Number of Times Ordered"), 
    align = "lcr"
  )
```

```{r, include = FALSE}
#live session
  instacart %>%
    filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
    group_by(aisle) %>% 
    count(product_name) %>%
    mutate(rank = min_rank(desc(n))) %>% 
    filter(rank <= 3) %>% 
    arrange(aisle, rank) %>% 
    knitr::kable(
      col.names = c("Aisle", "Product Name", "Number of Times Ordered", "Rank"), 
      align = "lccr"
    )
```

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r, message = FALSE}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  select(product_name, order_dow, order_hour_of_day) %>%
  group_by(product_name, order_dow) %>%
  summarise(
    mean_hour = mean(order_hour_of_day)
  ) %>%
  mutate(
    order_dow = case_when(
    order_dow == 0 ~ "Sunday",
    order_dow == 1 ~ "Monday",
    order_dow == 2 ~ "Tuesday",
    order_dow == 3 ~ "Wednesday",
    order_dow == 4 ~ "Thursday",
    order_dow == 5 ~ "Friday",
    order_dow == 6 ~ "Saturday",
    TRUE ~ "")
  ) %>%
  rename("Product Name" = product_name) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>%
  knitr::kable()
```

```{r, include = FALSE}
#live session
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

### Problem 2

**Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).**

```{r}
accel_df = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>%
  group_by(week) %>%
  pivot_longer(
    cols = activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity"
  ) %>%
  mutate(
    day = factor(day),
    day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
    weekend = factor(case_when(
      day == "Saturday" ~ "Weekend",
      day == "Sunday" ~ "Weekend",
      TRUE ~ "Weekday"
    )),
    minute = as.numeric(minute)
  ) %>% 
  select(week, day, weekend, day_id, minute, activity)
```

The _accel_df_ dataset contains data on accelerometer activity collected on a 63 year-old male with BMI 25. Each row in the dataset is the activity in one minute in a day within a week. The dataset has `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. 

**Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?**

```{r}
accel_df %>% 
  group_by(week, day) %>% 
  summarise(
    activity_day = sum(activity)
  ) %>% 
  knitr::kable()
```

In the first two weeks, he was less active on mondays and progressivly increased his activity until friday.
He tended to be fairly active during weekends, except for the last two weeks - he probably did not use his accelerometer on the last two saturdays.

**Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.**

```{r}
accel_df %>% 
  ggplot(aes(x = minute, y = activity, color = day))+
  geom_line() + 
  labs(
    title = "Activity of Mr. X",
    x = "Minute of the Day",
    y = "Activity",
      caption = "63 year-old, male, BMI 25"
  ) + 
  scale_x_continuous(
    limits = c(0, 1440),
    breaks = c(0, 360, 720, 1080, 1440),
    labels = c("00:00", "06:00", "12:00", "18:00", "00:00")
  )
```

He starts his activities around 5h and diminishes it around 22h. On sundays, the patient is more active around noon while during other days of the week he is more active at night. There is somewhat of a peak in activity around 17-18h on weekends apparently.

### Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```

*Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.**

The _my_noaa_ dataset contains data for all New York state weather stations from January 1, 1981 through December 31, 2010 acquired from the `rnoaa` package. The dataset has `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns, and it contains daily information on precipitation, snowfall, snow depth, and maximum and minimum temperatures. The dataset contains `r (ny_noaa %>% nrow()) - (ny_noaa %>% drop_na() %>% nrow())` days(`r round((nrow(ny_noaa) - (ny_noaa %>% drop_na() %>% nrow()))/(nrow(ny_noaa))*100, digits = 2)`%) with missing observations.

**Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units.**

```{r}
ny_noaa_clean =
ny_noaa %>%
  mutate(
    prcp = prcp/10,
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    tmax = tmax/10,
    tmin = tmin/10
    ) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-", convert = TRUE)
```

**For snowfall, what are the most commonly observed values? Why?**

```{r}
ny_noaa_clean %>% 
  count(snow) %>% 
  arrange(desc(n))
```

The most common observed values for snow in the dataset is 0 and NA. This happens because since _snow_ can take any value from 0 upwards, and in NY there are more days witout snow, it will be right skewed with many days with 0mm of snow. Also, there is a lot of missing data in this dataset, all coded the same.

**Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?**

```{r}
ny_noaa_clean %>%
  drop_na(tmax) %>% 
  group_by(id, year, month) %>% 
  mutate(
    mean_tmax = mean(tmax),
    month = month.name[month]
  ) %>%
  filter(month %in% c("January", "July")) %>%
  ggplot(aes(x = year, y = mean_tmax)) +
  geom_hex() +
  facet_grid(.~month) + 
  labs(
    x = "Year",
    y = "Mean Maximum daily Temperature (ºC)",
    fill = "# of NYS Weather Stations"
  )
```


**Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.**

```{r, include = FALSE}
#live session

#3 plot1 - contour/ hex (not scatter)  plot2 ridge  patchwork
```

