---
title: "Homework 3"
author: "Thiago de Araujo - UNI tbd2117"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

Loading Intracart data.

```{r}
library(p8105.datasets)
data("instacart")
```

**Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.**

```{r, message = FALSE}
order_df=
instacart %>% 
  group_by(order_id) %>% 
  summarize(
    n_item = n(),
    n_reorder = sum(reordered)
  ) %>% 
  mutate(reorder_pct = n_reorder / n_item)
```

The _instacart_ dataset constains data on products ordered through a online grocery service that allows you to shop online from local stores. Each row in the dataset is a product from an order (there is a single order per user in this dataset). The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns. It contains products from `r n_distinct(pull(instacart, order_id))` orders with an average `r round(mean(pull(order_df, n_item)), digits = 2)` items per order. In each order, on average, `r round(mean(pull(order_df, reorder_pct))*100, digits = 2)`% of items are reordered.

**How many aisles are there, and which aisles are the most items ordered from?**

```{r, message = FALSE}
aisle_df =
  instacart %>% 
    group_by(aisle) %>% 
    summarize(
      n_item = n()
    ) %>%
    arrange(desc(n_item))
```

There are `r nrow(aisle_df)` aisles with most itens ordered from the following aisles: `r head(pull(aisle_df, aisle))`.

**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r, message = FALSE}
aisle_df %>% 
  filter(n_item > 10000) %>%
  ggplot(aes(x = reorder(aisle, desc(n_item)), y = n_item)) + 
  geom_col() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle=45, hjust = 1)
  ) + 
  labs(
    title = "Number of Items Ordered in each Aisle",
    x = "Aisle",
    y = "Number of Items Ordered",
    caption = "Limited by aisles with more than 10,000 items orderd"
  )
```

**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r, message = FALSE}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  select(aisle, product_name) %>%
  group_by(aisle, product_name) %>%
  summarise(
    n_item = n()
  ) %>%
  arrange(desc(n_item)) %>%
  slice(1:3) %>% 
  mutate(
    aisle = str_to_title(aisle)
  ) %>% 
  knitr::kable(
    col.names = c("Aisle", "Product Name", "Number of Times Ordered"), 
    align = "lcr"
  )
```

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r, message = FALSE}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  select(product_name, order_dow, order_hour_of_day) %>%
  group_by(product_name, order_dow) %>%
  summarise(
    mean_hour = mean(order_hour_of_day)
  ) %>%
  mutate(
    order_dow = case_when(
    order_dow == 0 ~ "Sunday",
    order_dow == 1 ~ "Monday",
    order_dow == 2 ~ "Tuesday",
    order_dow == 3 ~ "Wednesday",
    order_dow == 4 ~ "Thursday",
    order_dow == 5 ~ "Friday",
    order_dow == 6 ~ "Saturday",
    TRUE ~ "")
  ) %>%
  rename("Product Name" = product_name) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>%
  knitr::kable()
```

### Problem 2